\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{tensor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{abstract}
\usepackage{titlesec}
\usepackage{cite}

% Define operators
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\dist}{dist}

% Theorems and definitions
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{property}{Property}

% Title formatting
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}

% Spacing
\setlength{\parskip}{1em}
\setlength{\parindent}{0em}
\setlength{\absleftindent}{0mm}
\setlength{\absrightindent}{0mm}

\title{\vspace{-2cm}\textbf{Narrative Forumlae}}
\author{\textbf{Paul Lowndes} \\ \href{mailto:ZeroTrust@NSHkr.com}{\texttt{ZeroTrust@NSHkr.com}}}
\date{\small January 4, 2025}

\begin{document}

\maketitle
\vspace{-1.5em}

\begin{abstract}

\end{abstract}


\section{Definitions}

\begin{itemize}
    \item \( M \): Differentiable manifold representing the narrative space.
    \item \( g \): Metric tensor on \( M \).
    \item \( c \in M \): A concept embedding, a point on the manifold.
    \item \( \gamma: [0,1] \rightarrow M \): A smooth, parameterized path on the manifold, representing a narrative.
    \item \( \gamma_T \): Geodesic path representing the "ground truth" narrative.
    \item \( LCM_i \): Large Concept Model of agent \( a_i \), which generates narrative paths.
	\item \( a_i \): Agent \( i \).
    \item \( A_i \): Tensor parameters encoding knowledge, beliefs, and influence, specific to agent \( a_i \) within the LCM.
	\item \( I \): Influence tensor field over \( M \).
    \item \( V: M \rightarrow [0, 1] \): Veracity function.
    \item \( T \subset M \): "Ground truth" region on the manifold.
	
\end{itemize}

\section{Narrative Dynamics}

\begin{definition}[Narrative Path]
A narrative \( N_{i,t} \) of agent \( a_i \) at time \( t \) is represented as a path \( \gamma_{i,t} \) on the manifold \( M \):
\begin{equation}
    \gamma_{i,t}(s) = (c_{i,1}, c_{i,2}, \dots, c_{i,T})
\end{equation}
where \( \gamma_{i,t}(0) = c_{i,1}, \gamma_{i,t}(1/T) = c_{i,2}, \dots, \gamma_{i,t}(1) = c_{i,T} \), and \( c_{i,k} \) are concept embeddings generated by \( LCM_i \).
\end{definition}

\begin{definition}[Narrative Generation by LCM]
The next concept embedding \( c_{i,k+1} \) in the narrative path is generated by the LCM as follows:
\begin{equation}
    c_{i,k+1} = LCM_i(c_{i,1:k}, K_{i,t}, B_{i,t}, C_t, A_i)
\end{equation}
where \( c_{i,1:k} \) is the sequence of preceding concept embeddings, \( K_{i,t} \) is the agent's knowledge set, \( B_{i,t} \) is the agent's belief set, \( C_t \) is the context at time \( t \), and \( A_i \) are agent-specific parameters.
\end{definition}

\section{Deviation from Ground Truth}

\begin{definition}[Geodesic Deviation Score (GDS)]
The Geodesic Deviation Score measures the deviation of a narrative path \( \gamma_{i,t} \) from the "ground truth" geodesic \( \gamma_T \):
\begin{equation}
    GDS(\gamma_{i,t}, \gamma_T) = \int_0^1 \| \dot{\gamma}_{i,t}(s) - \dot{\gamma}_T(s) \| ds
\end{equation}
where \( \dot{\gamma} \) represents the tangent vector of the path.
\end{definition}

\begin{definition}[Torsion]
The torsion of a narrative path \( \gamma_{i,t} \) is a measure of its "twisting":
\begin{equation}
    \text{Torsion}(\gamma_{i,t}) = \int_0^1 | \tau(s) | ds
\end{equation}
where \( \tau(s) \) is the torsion of the path at parameter \( s \).
\end{definition}

\section{Veracity Function}

The veracity function \( V \) can be redefined to incorporate geometric measures:

\begin{equation}
V(\gamma_{i,t}) = w_1 \cdot (1 - GDS(\gamma_{i,t}, \gamma_T)) + w_2 \cdot (1 - \text{Torsion}(\gamma_{i,t})) + \dots
\end{equation}

where \( w_1, w_2, \dots \) are weights assigned to each factor.


\section{Narrative Dynamics}

\begin{definition}[Narrative Path]
A narrative \( N_{i,t} \) of agent \( a_i \) at time \( t \) is a path \( \gamma_{i,t} \) on the manifold \( M \).
\end{definition}

\begin{definition}[Narrative Evolution]
The evolution of a narrative path is governed by a flow on the manifold:
\begin{equation}
\frac{dN_{i,t}}{dt} = F(N_{i,t}, A_i, I(N_{i,t}))
\end{equation}
where \( F \) is a function that describes the dynamics of the narrative flow, influenced by the agent's internal parameters \( A_i \) and the influence tensor field \( I \) at the current narrative point.
\end{definition}

\section{Influence Tensor Field}

The influence tensor field \( I \) can be defined at each point \( c \in M \) to represent the potential for influence or manipulation:

\begin{equation}
I(c) = \sum_{j \neq i} \alpha_{ij} \cdot G_{j(c)}
\end{equation}

where:
\begin{itemize}
    \item \( \alpha_{ij} \): Influence weight of agent \( a_j \).
    \item \( G_{j(c)} \): Influence tensor of agent \( a_j \) at point \( c \), encoding the magnitude, direction, and source of influence.
\end{itemize}

\section{Veracity and Geometry}

\begin{equation}
V(\gamma_{i,t}) = w_1 \cdot (1 - GDS(\gamma_{i,t}, \gamma_T)) + w_2 \cdot (1 - \text{Torsion}(\gamma_{i,t})) + \dots
\end{equation}

where:
\begin{itemize}
    \item \( GDS(\gamma_{i,t}, \gamma_T) = \int_0^1 \| \dot{\gamma}_{i,t}(s) - \dot{\gamma}_T(s) \| ds \) (Geodesic Deviation Score)
    \item \( \text{Torsion}(\gamma_{i,t}) = \int_0^1 | \tau(s) | ds \) (\( \tau(s) \) is the torsion of the path)
\end{itemize}


\end{document}
