\documentclass[12pt, a4paper]{article}

% --- PACKAGES ---
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{enumitem}
\usepackage{abstract}
\usepackage{titlesec}
\usepackage{float}
\usepackage{amsthm} 

\usepackage{natbib}  
 



\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\setlength{\parskip}{0.7em}
\setlength{\parindent}{0em}
\renewcommand{\abstractname}{\vspace{-\baselineskip}} % Remove "Abstract" title
\newtheorem{definition}{Definition}[section]

% --- DOCUMENT INFORMATION ---
\title{\textbf{CNS: A Practical Blueprint for Chiral Narrative Synthesis}}
\author{
    Paul Lowndes \\
    \small Conceptual AI Laboratory \\
    \small \href{mailto:ZeroTrust@NSHkr.com}{\texttt{ZeroTrust@NSHkr.com}}
}
\date{July 29, 2025}

\begin{document}

\maketitle
\vspace{-2em}

\begin{abstract}
% REVISION: Abstract improved for clarity and concision
Knowledge synthesis from conflicting sources remains a fundamental challenge in artificial intelligence. We present Chiral Narrative Synthesis (CNS), a practical engineering blueprint that transforms conflicting information into coherent knowledge through multi-agent dialectical reasoning. Our framework introduces four key innovations: (1) Structured Narrative Objects (SNOs) that replace simple vectors with rich representations combining hypotheses, reasoning graphs, evidence sets, and trust scores; (2) a transparent multi-component critic pipeline replacing black-box evaluation with specialized assessors for grounding, logic, and novelty; (3) LLM-powered generative synthesis that transcends naive averaging through dialectical reasoning; and (4) "Evidential Entanglement," a novel metric identifying productive conflicts between narratives arguing over shared data. We provide both the system architecture and a concrete research roadmap addressing critical implementation challenges—from narrative ingestion to model development—establishing a foundation for automated knowledge discovery systems capable of reconciling contradictory information into robust insights.
\end{abstract}

\section{Introduction}
Complex domains—from scientific research to intelligence analysis—require synthesizing incomplete, uncertain, and contradictory information into coherent knowledge. Despite AI's success in pattern recognition, the cognitive challenge of reconciling conflicting hypotheses remains unsolved \citep{Boström2017}. This challenge stems from argumentation's inherent complexity: claims exist within intricate webs of evidence and reasoning that resist simple computational approaches \citep{Toulmin2003}. 

We propose Chiral Narrative Synthesis (CNS), a computational framework that operationalizes knowledge synthesis by treating hypotheses as mathematically evaluable data structures rather than simple text. Moving beyond conceptual models to practical implementation, CNS introduces four key advances:
\begin{enumerate}
    \item \textbf{Structured Narrative Objects (SNOs):} Rich data structures capturing hypotheses, logical reasoning graphs, evidence sets, and trust scores
    \item \textbf{Multi-Component Critic Pipeline:} Transparent evaluation replacing black-box oracles with specialized assessors for grounding, logic, and novelty
    \item \textbf{Generative Synthesis Engine:} LLM-powered dialectical reasoning that transcends naive vector averaging
    \item \textbf{Evidential Entanglement Metric:} Novel measure identifying narratives that oppose each other while arguing over shared evidence
\end{enumerate}

We formalize the dialectical process of resolving conflict and integrating independent knowledge. As such, CNS offers a promising computational approach to automated, robust, and auditable knowledge discovery.

% REVISION: Improved section introduction with clearer structure
\section{The CNS Architecture}
The CNS framework rests on three foundational pillars: (i) narrative representation through structured objects, (ii) multi-faceted evaluation via specialized critics, and (iii) synthesis through generative dialectical reasoning. We detail each component below.

% REVISION: Enhanced SNO section with clearer motivation and notation
\subsection{Structured Narrative Objects (SNOs)}
Traditional vector representations lose critical structural and evidential information necessary for dialectical reasoning. We address this limitation through Structured Narrative Objects (SNOs), which preserve the full richness of argumentative structure.

\begin{definition}[Structured Narrative Object]
% REVISION: Improved mathematical notation and clarity
An SNO is a 4-tuple $\mathcal{S} = (H, G, \mathcal{E}, T)$ where:
\begin{itemize}
    \item \textbf{Hypothesis Embedding} $H \in \mathbb{R}^d$: A $d$-dimensional dense vector encoding the narrative's central claim, enabling geometric similarity computations while preserving semantic content
    \item \textbf{Reasoning Graph} $G = (V, E_G)$: A directed acyclic graph with vertices $V$ representing sub-claims and edges $E_G \subseteq V \times V \times \mathcal{R}$ encoding typed logical relationships (e.g., "supports," "contradicts," "implies") from the relation set $\mathcal{R}$
    \item \textbf{Evidence Set} $\mathcal{E} = \{e_1, e_2, \ldots, e_n\}$: Pointers to grounding data sources, including document identifiers, data hashes, or persistent identifiers (DOIs), establishing verifiable connections to primary sources
    \item \textbf{Trust Score} $T \in [0, 1]$: A derived confidence measure computed by the critic pipeline, not an intrinsic property of the narrative
\end{itemize}
\end{definition}

% REVISION: Added transition and improved section clarity
This structured representation enables sophisticated reasoning operations while maintaining computational tractability through the vector embedding $H$.

\subsection{Multi-Component Critic Pipeline}
% REVISION: Clearer problem statement and solution
Traditional "oracle" critics suffer from opacity and unverifiability. We decompose evaluation into a transparent pipeline of specialized critics, each assessing distinct aspects of narrative quality. The final trust score $T$ and reward signal emerge from a weighted combination:

\begin{equation}
\text{Reward}(\mathcal{S}) = \sum_{i \in \{G, L, N\}} w_i \cdot \text{Score}_i(\mathcal{S})
\label{eq:reward}
\end{equation}

where $w_i$ are dynamically adjustable weights and the component scores are:
% REVISION: Enhanced critic descriptions with mathematical formulations
\begin{itemize}
    \item \textbf{Grounding Critic} $(\text{Score}_G)$: Evaluates evidential support by computing plausibility scores between claims in $G$ and evidence in $\mathcal{E}$ using a fine-tuned NLI model. For each vertex $v \in V$ with associated evidence $e \in \mathcal{E}$, we compute $p(v|e)$ and aggregate: $\text{Score}_G = \frac{1}{|V|}\sum_{v \in V} \max_{e \in \mathcal{E}} p(v|e)$
    
    \item \textbf{Logic Critic} $(\text{Score}_L)$: Assesses structural coherence of the reasoning graph $G$ using a Graph Neural Network trained to detect logical weaknesses (circular dependencies, orphaned claims, excessive branching). The GNN produces: $\text{Score}_L = f_{\text{GNN}}(G; \theta)$ where $\theta$ are learned parameters
    
    \item \textbf{Novelty-Parsimony Critic} $(\text{Score}_N)$: Balances innovation against redundancy and complexity. Given existing high-trust SNOs $\{\mathcal{S}_i\}_{i=1}^M$, compute: $\text{Score}_N = \alpha \cdot \min_i \|H - H_i\|_2 - \beta \cdot \frac{|E_G|}{|V|}$ where the first term rewards novelty and the second penalizes graph complexity
\end{itemize}
% REVISION: Improved transition and clarity
Dynamic weight adjustment enables context-sensitive evaluation—prioritizing grounding in empirical domains, logic in theoretical contexts, or novelty during exploratory phases.

\subsection{Generative Synthesis Engine}
% REVISION: Stronger motivation and clearer description
Vector averaging fails to capture the nuanced reasoning required for genuine knowledge synthesis. Our Generative Synthesis Engine employs an LLM fine-tuned for dialectical reasoning to produce semantically coherent resolutions of conflicting narratives.

The synthesis workflow operates as follows:
% REVISION: Improved synthesis process description
\begin{enumerate}
    \item \textbf{Chiral Pair Selection:} Identify SNO pairs $(\mathcal{S}_A, \mathcal{S}_B)$ with high chirality and evidential entanglement scores (detailed in Section 3.2)
    
    \item \textbf{Dialectical Prompt Construction:} Transform SNOs into a structured prompt that preserves argumentative structure:
        \begin{quote}
        \texttt{NARRATIVE\_A}: $\{H_A, G_A, \mathcal{E}_A\}$ \\
        \texttt{NARRATIVE\_B}: $\{H_B, G_B, \mathcal{E}_B\}$ \\
        \texttt{CONFLICT\_ANALYSIS}: Identify contradictions in hypotheses while preserving shared evidence \\
        \texttt{SYNTHESIS\_TASK}: Generate $\mathcal{S}_C$ that resolves conflicts through higher-order reasoning
        \end{quote}
    
    \item \textbf{Candidate Generation:} The LLM produces a new SNO $\mathcal{S}_C = (H_C, G_C, \mathcal{E}_C, \varnothing)$ where $\mathcal{E}_C \supseteq \mathcal{E}_A \cap \mathcal{E}_B$ and the trust score remains unassigned pending evaluation
    
    \item \textbf{Critic Evaluation:} The candidate $\mathcal{S}_C$ enters the critic pipeline to determine its viability
\end{enumerate}
This approach models synthesis not as a mathematical blend, but as an act of creative, reasoned generation.

\section{System Dynamics and Workflow}
The full CNS system operates in a continuous loop, driven by precise metrics and specialized agent actions.

\subsection{The Narrative Ingestion Pipeline: A Key Research Challenge}
A critical prerequisite for the CNS ecosystem is the ability to generate SNOs from unstructured source materials (e.g., academic papers, intelligence reports). This process, a form of advanced argumentation mining \citet{Lippi2016ArgMining}, is a major research challenge in itself. Our proposed initial pipeline is as follows:
\begin{enumerate}
    \item \textbf{Hypothesis Extraction:} An LLM is prompted to read a source document and output a concise summary of its central claim or hypothesis. This summary is then embedded to produce the initial `Hypothesis Embedding (H)`.
    \item \textbf{Reasoning Graph Construction:} We will explore a hybrid approach. First, use an LLM to identify key sub-claims and their relationships (e.g., "premise A supports conclusion B"). Then, formalize these extracted relationships into the directed graph structure of `G`. The development of robust prompts and validation techniques for this step is a primary research task.
    \item \textbf{Evidence Set Population:} Use a combination of pattern matching (for explicit citations like DOIs) and semantic search to link claims within the `Reasoning Graph (G)` to specific sentences or data points in the source document, which then form the `Evidence Set (E)`.
\end{enumerate}
This pipeline represents a core workstream of the project, turning a critical dependency into a defined research objective.


\subsection{Refined Relational Metrics}
The concept of "chirality" is made more precise by distinguishing between opposition and shared context. This allows the system to identify the most productive conflicts.

\begin{definition}[Chirality Score]
The Chirality Score remains a useful measure of opposing \textit{hypotheses}. It is calculated using the Hypothesis Embeddings ($H$) from two SNOs:
\[
\text{CScore}(SNO_i, SNO_j) = (1 - H_i \cdot H_j) \cdot (T_i \cdot T_j)
\]
This score is high when two well-supported narratives propose contradictory central claims.
\end{definition}

\begin{definition}[Evidential Entanglement]
This new metric measures the degree to which two narratives are arguing over the same data. It is calculated using the Jaccard similarity of their \textit{Evidence Sets (E)}:
\[
\text{EScore}(SNO_i, SNO_j) = \frac{|E_{\text{set}, i} \cap E_{\text{set}, j}|}{|E_{\text{set}, i} \cup E_{\text{set}, j}|}
\]
\end{definition}

\textbf{Synthesis Trigger:} The synthesis process is prioritized for pairs with \textbf{both high Chirality and high Entanglement}. These represent two well-supported, opposing theories that are attempting to explain the same set of facts—the most fertile ground for a novel synthesis.

\subsection{System Operational Loop}
The full system operates as follows:
\begin{enumerate}
    \item \textbf{Population:} The system maintains a dynamic population of SNOs, initially populated via the Narrative Ingestion Pipeline.
    \item \textbf{Relational Mapping:} The system continuously computes relational scores. To ensure scalability, this is a two-step process. First, an Approximate Nearest Neighbor index (e.g., LSH \citep{Indyk1998LSH}) on the $H$ vectors is used to efficiently pre-filter a small set of candidate pairs with high potential `CScore`. Second, the more computationally intensive `EScore` is calculated only for these pre-filtered pairs.
    \item \textbf{Agent Action:}
        \begin{itemize}
            \item \textbf{Synthesizer Agents} select high-chirality, high-entanglement pairs and pass them to the \textbf{Generative Synthesis Agent (LLM)} to create new candidate SNOs.
            \item \textbf{Narrator Agents} can still perform exploration or refinement, for example, through the guided exploration method described below.
        \end{itemize}
    \item \textbf{Evaluation:} All newly generated SNOs are rigorously evaluated by the \textbf{Multi-Component Critic} pipeline to determine their Trust Score $T$.
    \item \textbf{Selection:} SNOs that achieve a high Trust Score are integrated into the main population. Low-scoring SNOs are archived. This constitutes the survival-of-the-fittest mechanism for knowledge.
\end{enumerate}

\subsection{Guided Narrative Exploration via Latent Space Targeting}
Instead of directly modifying an SNO's components via gradient ascent, which can lead to internal inconsistency, we propose a more robust generative method for narrative exploration. When an agent seeks to refine an SNO$_i$ that is part of a chiral pair with SNO$_j$, it can compute a *target embedding* in a novel region of the conceptual space.

The target embedding, $H_{\text{target}}$, can be calculated as:
\begin{equation} \label{eq:crga_reworked}
H_{\text{target}} = H_{i} + \alpha \nabla_{H_i} \text{Reward}(SNO_i) + \beta \cdot \text{CScore}(SNO_i, SNO_j) \frac{H_{i} - H_{j}}{\|H_{i} - H_{j}\|}
\end{equation}
This vector represents a conceptual direction that is rewarded by the critic while also being repelled from its chiral partner. This $H_{\text{target}}$ is not used to modify SNO$_i$. Instead, it is used to prompt a generative agent: "Generate a new SNO whose core hypothesis is semantically close to $H_{\text{target}}$, drawing inspiration from the reasoning and evidence of SNO$_i$." This prompts the creation of a new, fully-formed candidate SNO that explores the space between existing ideas, which can then be evaluated by the critic pipeline.


\section{Discussion and Future Work}
This CNS blueprint creates a plausible and powerful system by making the abstract components of earlier models concrete. It directly addresses key philosophical and practical challenges.

\textbf{On the Nature of "Truth":} The system avoids the "Truth Oracle" problem. "Truth" is not a predefined target but an emergent property, represented by regions of the state space containing diverse, coherent, and highly explanatory SNOs. This aligns with a Kuhnian view of scientific truth as a provisional, ever-improving model of reality \citep{Kuhn1962}.

\textbf{Interpretability and Grounding:} The framework is inherently more interpretable. The success of a given SNO is not a mystery; it can be explained by its individual scores from the critic pipeline (e.g., "This narrative is trusted because its logic is sound and its evidence is verifiable, despite being similar to existing ideas"). The `Evidence Set (E)` and `Grounding Critic` directly solve the grounding problem, anchoring the abstract narrative space to verifiable data.

\textbf{Future Work and Research Roadmap:} The primary challenge shifts from conceptual design to engineering, tuning, and evaluation. This proposal defines the following key research thrusts:

\begin{enumerate}
    \item \textbf{Development of Critic Models:} A significant effort will be dedicated to developing the GNN for the Logic Critic and the NLI model for the Grounding Critic. This involves curating specialized datasets and defining appropriate model architectures for assessing structural integrity and evidential plausibility.
    
    \item \textbf{Bootstrapping the Generative Synthesizer:} The quality of the Generative Synthesis Agent is dependent on its training. We propose a multi-stage strategy:
        \begin{itemize}
            \item \textbf{Phase 1 (Few-Shot Prompting):} Initially, the system will rely on the rich, structured prompts enabled by the SNO format.
            % FIX START: Replaced backticks `...` with \texttt{...} and math mode to fix fatal compiler errors.
            \item \textbf{Phase 2 (Self-Improvement):} The CNS system itself will generate training data. High-scoring syntheses \texttt{SNO\_C} generated from pairs (\texttt{SNO\_A}, \texttt{SNO\_B}) will be archived as positive training examples $(A, B) \to C$, creating a flywheel for continuous improvement.
            % FIX END
            \item \textbf{Phase 3 (Human-in-the-Loop):} We will develop an interface for human experts to review, rate, and correct syntheses, providing a gold-standard dataset for fine-tuning the LLM on high-quality dialectical reasoning.
        \end{itemize}

    \item \textbf{Formal Evaluation Protocol:} To measure the system's success, we will develop a formal evaluation protocol. A candidate experiment involves seeding the system with SNOs derived from papers representing historical scientific debates (e.g., plate tectonics vs. geosyncline theory). The primary success metric will be the system's ability to generate a synthesized SNO that aligns with the modern scientific consensus, evaluated both by its Critic Score and by human expert review.
\end{enumerate}


\section{Conclusion}
Chiral Narrative Synthesis provides a comprehensive blueprint for a multi-agent system capable of automated knowledge discovery. By integrating a rich narrative structure (SNO), a transparent evaluation pipeline (Multi-Component Critic), a sophisticated generative engine (LLM Synthesizer), and precise relational metrics (Chirality and Entanglement), this framework moves beyond a purely conceptual model. It lays out a practical path and a clear research roadmap toward building AI systems that can reason about, reconcile, and synthesize conflicting information to generate novel and robust insights.

% REVISION: Enhanced bibliography formatting and consistency
\bibliographystyle{plainnat}
\bibliography{references.bib}

\end{document}